DATA 624 Homework 6

### 2020-04-05 DATA624/HW7.Rmd
More work on HW 7

---
title: "DATA 624 Homework 6"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(kableExtra)
library(glmnet)
library(pls)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)
# Make this reproducible
set.seed(42)
```

@@

```{r}
df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>%
  mutate(y = permeability)
in_train <- createDataPartition(df$y, times = 1, p = .8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]

@@

# Plot model RMSE vs different values of components
plot(pls_model, main=paste("Training Set RMSE Minimized at", pls_model$bestTune$ncomp, "Components"))
```

@@

plot(pcr_model, main = paste("Training Set RMSE Minimized at", pcr_model$bestTune, "Components"))

@@

```{r}
x <- model.matrix(y ~ ., data=train_df)
x_test <- model.matrix(y ~ ., data=test_df)
rr_cv <- cv.glmnet(x, train_df$y, alpha = 0)
rr_model <- glmnet(x, train_df$y, alpha = 0, lambda = rr_cv$lambda.min)
rr_predictions <- as.vector(predict(rr_model, x_test))
rr_results <- data.frame(Model = "Ridge Regression", RMSE = caret::RMSE(rr_predictions, test_df$y), Rsquared = caret::R2(rr_predictions, test_df$y)) 
rr_results %>%
  kable() %>%
  kable_styling()
```

@@

@@

@@

(A) Start `R` and use these commands to laod the data:

```{r}
library(AppliedPredictiveModeling)
data(chemicalManufacturing)
```

@@

(B) A small percentage of cells in the predictor set contain missing values.  Use an imputation function to fill in these missing values (e.g., see Sect. 3.8).

(C) Split the data into a training and a test set, pre-process the data, and tune a model of your choice from this chapter.  What is the optimal value of the performance metric?

(D) Predict the response for the test set.  What is the value of the performance metric and how does this compare with the resampled performance metric on the training set?

(E) Which predictors are most important in the model you have trained?  Do either the biological or process predictors dominate the list?

(F) Explore the relationships between each of the top predictors and the response.  How could this information be helpful in improving yield in future runds of the manufacturing process?

>>>>

---
title: "DATA 624 Homework 6"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 4
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, 
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(kableExtra)
library(glmnet)
library(pls)
library(corrplot)
library(tidyr)
library(RColorBrewer )
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)
```

@@

```{r}
df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>%
  mutate(y = permeability)

# Make this reproducible
set.seed(42)

in_train <- createDataPartition(df$y, times = 1, p = 0.8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]

@@

# Plot model RMSE vs different values of components
title <- paste("Training Set RMSE Minimized at",
               pls_model$bestTune$ncomp,
               "Components")
plot(pls_model, main = title)
```

@@

title <- paste("Training Set RMSE Minimized at",
               pcr_model$bestTune,
               "Components")
plot(pcr_model, main = title)

@@

```{r}
x <- model.matrix(y ~ ., data = train_df)
x_test <- model.matrix(y ~ ., data = test_df)
rr_cv <- cv.glmnet(x, train_df$y, alpha = 0)
rr_model <- glmnet(x, train_df$y, alpha = 0, lambda = rr_cv$lambda.min)
rr_predictions <- as.vector(predict(rr_model, x_test))
rr_results <- data.frame(Model = "Ridge Regression",
                         RMSE = caret::RMSE(rr_predictions, test_df$y),
                         Rsquared = caret::R2(rr_predictions, test_df$y))
rr_results %>%
  kable() %>%
  kable_styling()
```

@@

**None of the models did a better job (had a higher $R^2$) than the PLS model.**

@@

**No, because the best model's (PLS) $R^2$ was really low.  The model did not have much explanatory power.**

@@

```{r}
library(AppliedPredictiveModeling)
data(ChemicalManufacturingProcess)
```

@@

A small percentage of cells in the predictor set contain missing values.  Use an imputation function to fill in these missing values (e.g., see Sect. 3.8).

**I will use KNN to impute values.  The `caret` package makes it simple:**

```{r}
# Make this reproducible
set.seed(42)

knn_model <- preProcess(ChemicalManufacturingProcess, "knnImpute")
df <- predict(knn_model, ChemicalManufacturingProcess)
```

Split the data into a training and a test set, pre-process the data, and tune a model of your choice from this chapter.  What is the optimal value of the performance metric?

**I will pre-process the data removing the zero variance variables prior to splitting the data.**

```{r}
df <- df %>%
  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))
```

**Now the data will be split into training and test set using an 80:20 split.**

```{r}
in_train <- createDataPartition(df$Yield, times = 1, p = 0.8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]
```

**Given the preformance of the PLS in the previous exercise, I will use it again.  I'm going to further pre-process by centering and scaling the predictors.**

```{r}
pls_model <- train(
  Yield ~ ., data = train_df, method = "pls",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

# Plot model RMSE vs different values of components
title <- paste("Training Set RMSE Minimized at",
               pls_model$bestTune$ncomp,
               "Components")
plot(pls_model, main = title)
```

**For the performance metric I will again look at the RMSE and the $R^2$.  The optimal values for RMSE would be zero, and the optimal $R^2$ would be 1.  Here's the performance metrics for the optimal model on the training set.**

Predict the response for the test set.  What is the value of the performance metric and how does this compare with the resampled performance metric on the training set?

```{r}
# Make predictions
pls_predictions <- predict(pls_model, test_df)
# Model performance metrics
results <- data.frame(RMSE = caret::RMSE(pls_predictions, test_df$Yield),
           Rsquared = caret::R2(pls_predictions, test_df$Yield))
results %>%
  kable() %>%
  kable_styling()
```

**We have an RMSE of about `r round(results$RMSE, 2)` and a $R^2$ of roughly `r round(results$Rsquared, 2)` on the test set.**

Which predictors are most important in the model you have trained?  Do either the biological or process predictors dominate the list?

```{r}
varImp(pls_model) %>%
  plot(, top = 20, main = "Important Variables (Cliped at 20)")
```

```{r}
pls_importance <- varImp(pls_model)$importance %>% 
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  arrange(desc(Overall)) %>%
  mutate(importance = row_number())

pls_importance %>%
  mutate(Variable = gsub("[0-9]+", "", Variable)) %>%
  group_by(Variable) %>%
  tally() %>%
  arrange(desc(n)) %>%
  kable() %>%
  kable_styling()
```

**The process predictors dominate the list.**

Explore the relationships between each of the top predictors and the response.  How could this information be helpful in improving yield in future rounds of the manufacturing process?

```{r}
df %>%
  select_at(vars(Yield, pls_importance$Variable)) %>%
  cor() %>%
  corrplot(type = "lower", order = "hclust", tl.col = "black", tl.srt = 45)
  #corrplot(., type = "upper", order = "hclust",
  #       col = brewer.pal(n = 8, name = "RdBu"))
```
  merge(pls_importance) %>%
  ggplot(aes(value, Yield, color = Variable)) +
  theme(legend.position = "none") +
  geom_point() %>%
  facet_wrap(~ Variable)
```
